
\chapter{Week12}

\section{Monday}\index{week6_Tuesday_lecture}
\subsection{Comments on Final Project}

The final project aims to solve the optimization problem given below:
\begin{equation}\label{Eq:12:1}
\begin{array}{ll}
\min&f(x,y)=\frac{1}{2}(\bm x\trans\bm A\bm x+\bm y\trans\bm A\bm y) - \bm b\trans\bm x-\bm c\trans\bm y:=q_b(\bm x)+q_c(\bm y)\\
\mbox{such that}&h_1(\bm x,\bm y):=\frac{1}{2}(\bm x\trans\bm x-1)=0\\
&h_2(\bm x,\bm y):=\frac{1}{2}(\bm y\trans\bm y-1)=0\\
&h_3(\bm x,\bm y):=\bm x\trans\bm y=0
\end{array}
\end{equation}

The Lagrangian function is given by:
\begin{equation}
L(\bm x,\bm y,\lambda)=q_b(\bm x)+q_c(\bm y)+\sum_{i=1}^3h_i(\bm x,\bm y)\lambda_i
\end{equation}

The stationary point should satisfy:
\begin{align*}
\nabla_{\bm x}L=
\bm{Ax}-\bm b+\lambda_1\bm x+\lambda_3\bm y&=0\\
\nabla_{\bm y}L=
\bm{Ay}-\bm c+\lambda_2\bm y+\lambda_3\bm x&=0\\
\|h(\bm x,\bm y)\|^2&=0
\end{align*}

Based on the stationarity condition, the stopping criteria is given by:
\[
\max\left\{
\frac{\|\nabla_{\bm x}L\|}{\|\bm b\|+1},
\frac{\|\nabla_{\bm y}L\|}{\|\bm c\|+1},
\|h(\bm x)\|
\right\}\le\mbox{tol}
\]

The problem(\ref{Eq:12:1}) admits its global minimum since the constraint set is compact. Three methods are suggests: ADMM (the lecturer used this one); ALMM; quadratic penalty method.

\paragraph{Suggested ADMM method}
The key of ADMM is to separate variables in constraint. Thus to eliminate $h_3(\bm x,\bm y)=0$, define the second-kind Lagrangian function
\[
\hat L(\bm x,\bm y,\lambda) = q_b(\bm x)+q_c(\bm y)+\lambda_3\bm x\trans\bm y+\frac{h}{2}(\bm x\trans\bm y)^2
\]

It suffices to do the minimization
\begin{equation}
\begin{array}{ll}
\min&\hat L(\bm x,\bm y,\lambda)\\
&\bm x\in X=\{\bm x\mid \bm x\trans\bm x=1\}\\
&\bm y\in Y=\{\bm y\mid \bm y\trans\bm y=1\}\\
\end{array}
\end{equation}

The update rule for $\lambda_3$ is therefore given by ($\tau=1.618$)
\[
\lambda_3=\lambda_3+\tau\rho(\bm x\trans\bm y)
\]

To implement such a problem, you need to develop a solver for the \emph{trust region} sub-problem
\begin{equation}\label{Eq:12:4}
\begin{array}{ll}
\min&\frac{1}{2}\bm p\trans\bm B\bm p+\bm g\trans\bm p\\
&
\bm p\trans\bm p=\Delta^*
\end{array}
\end{equation}

How to get the \emph{global minimum} for the non-convex problem (\ref{Eq:12:4})? Unfortunately, solving this problem is at least in polynomial time in general, which is not acceptable.


\subsection{Trust Region Method}
However, our $\bm B$ for (\ref{Eq:12:4}) has the special form
\[
\bm A+\lambda_1\bm I+\rho\bm y\bm y\trans
\]
with sparse $\bm A$ and dense rank 1 matrix $\bm y$. The key to solve (\ref{Eq:12:4}) is to get the inverse of $\bm B$. It suffices to apply the Sherman-Morrison Formula, or apply congugate gradient (command: pcg)

\paragraph{Another application of Trust Region Problem}

Our goal is to minimize $\min_{\bm x\in\mathbb{R}^n}f(\bm x)$. To choose the step-size and the direction, we apply the trust region method. First approximate $f(x)$ with quadratic problem:
\begin{align*}
f(x^*+p)&\approx
f(x^*)+\nabla\trans f(x^*)p+\frac{1}{2}p\trans\nabla^2f(x^*)p\\
&\approx
f(x^*)+\nabla\trans f(x^*)p+\frac{1}{2}p\trans\underbrace{\bm B}_{\text{Approximate Hessian}}p:=m(p)
\end{align*}
It suffices to choose the step size $p$ to minimize the quadratic function above. The constraint is that the $\|p\|$ should be small enough. Thus it suffices to solve
\begin{equation}\label{Eq:12:5}
\begin{array}{ll}
\min&\frac{1}{2}\bm p\trans\bm B\bm p+\bm g\trans\bm p\\
&\|\bm p\|\le\Delta
\end{array}
\end{equation}

The general trust region descent method is proposed below:
\begin{enumerate}
\item
Solve the TR subproblem (\ref{Eq:12:5}) to get $p$
\item
$\rho:=\frac{f(x) - f(x+p)}{m(0) - m(p)}$. 
\begin{enumerate}
\item
If $\rho<\frac{1}{4}$, decrease the trust region $\Delta$ to $\frac{1}{4}\Delta$.
\item
If $\rho>\frac{3}{4}$, increase $\Delta$ into $2\Delta$
\item
Else, keep $\Delta$.
\end{enumerate}
\item
\begin{enumerate}
\item
If $\rho\ge\frac{1}{4}$ and $\rho>\eta$, then $x\leftarrow x+p$
\item
Else, $x$ keeps unchanged.
\end{enumerate}
\end{enumerate}













