\section{Monday for MAT4002}\index{Monday_lecture}
\subsection{Continuous Functions on Compact Space}

\begin{proposition}\label{pro:5:3}
Let $f:X\to Y$ be continuous function on topological spaces, with $A\subseteq X$ compact. Then $f(A)\subseteq Y$ is compact.
\end{proposition}
\begin{proof}
Let $\{U_i\mid i\in I\}$ be an open cover of $f(A)$, i.e.,
\[
f(A)\subseteq\bigcup_{i\in I}U_i,\quad U_i\in\mathcal{T}_Y
\]
It follows that $\{f^{-1}(U_i)\mid i\in I\}$ is an open cover of $A$:
\[
A\subseteq f^{-1}\left(\bigcup_{i\in I}U_i\right)=\bigcup_{i\in I}f^{-1}(U_i)
\]
By the compactness of $A$, there exists finite subcover of $A$:
\[
A\subseteq\bigcup_{k=1}^nf^{-1}(U_{i_k}),
\]
which implies the constructed finite subcover of $f(A)$:
\begin{align*}
f(A)&\subseteq f(\cup_{k=1}^nf^{-1}(U_{i_k}))\\
&=\bigcup_{k=1}^nU_{i_k}
\end{align*}
\end{proof}

\begin{corollary}
\begin{enumerate}
\item
Suppose that $X$ is compact, and the mapping $f:X\to\mathbb{R}$ is continuous, then $f(X)$ is closed and bounded, i.e., there exists $m,M\in X$ such that $f(m)\le f(x)\le f(M),\ \forall x\in X$.
\item
Suppose moreover that $X$ is connected, then
\[
f(X)=[f(m),f(M)].
\]
\end{enumerate}
\end{corollary}

\begin{theorem}
The space $X,Y$ are compact iff $X\times Y$ is compact under product topology.
\end{theorem}
\begin{proof}
\begin{enumerate}
\item
\textit{Sufficiency:}
Given that $X\times Y$ is compact, consider the projection mapping (which is continuous):
\[
\left\{
\begin{aligned}
P_X&:X\times Y\to X\\
P_Y&:X\times Y\to Y
\end{aligned}
\right.
\]
By applying proposition~(\ref{pro:5:3}), $P_X(X\times Y)=X,P_Y(X\times Y)=Y$ are both compact.
\item
\textit{Necessity:}
Suppose that $\{W_i\}_{i\in I}$ is an open cover of $X\times Y$. Each open set $W_i$ can be written as:
\[
W_i = \bigcup_{j\in\mathcal{J}_i}U_{ij}\times V_{ij},\quad
U_{ij}\in\mathcal{T}_X,V_{ij}\in\mathcal{T}_Y.
\]
It follows that 
\[
X\times Y = \bigcup_{(i,j)\in K}U_{ij}\times V_{ij},\quad
K=\{(i,j)\mid i\in I, j\in\mathcal{J}_i\}
\]
Therefore, it suffices to show $\{U_{ij}\times V_{ij}\mid (i,j)\in K\}$ has a finite subcover of $X\times Y$.
\begin{itemize}
\item
Note that $X\times\{y\}\subseteq\bigcup_{(i,j)\in K}U_{ij}\times V_{ij}$ is compact for each $y\in Y$, which implies there exists finite $S_y\in K$ such that 
\[
X\times\{y\}\subseteq\bigcup_{s\in S_y}U_{s}\times V_{s}
\]
\item
w.l.o.g., assume that $y\in V_s,\forall s\in S_y$, since we can remove the $U_s\times V_s$ such that $y\notin V_s$.
Define the set $V_y:=\cap_{s\in S_y}V_s$, which is an open set containing $y$. We imply $\{V_y\}_{y\in Y}$ forms an open cover of $Y$.
By the compactness of $Y$, 
\[
\{V_{y_1},\dots,V_{y_m}\}
\]
forms a finite subcover of $Y$.
\item
For each $\ell=1,\dots,m$,
\[
X\times \{y_\ell\}\subseteq 
\bigcup_{s\in S_{y_\ell}}U_s\times V_s
\]
Note that for any $(x,y)\in X\times Y$, there exists $\ell\in\{1,\dots,m\}$ such that $y\in V_{y\ell}$, i.e., $y\in V_s$ for $\forall s\in S_{y\ell}$.
Therefore,
\[
X\times Y = \bigcup_{\ell=1}^m\bigcup_{s\in S_{y_\ell}}U_s\times V_s
\]
\end{itemize}
Now pick
\[
I'=\{i\in I\mid (i,j)\in\cup_{\ell=1}^m S_{y_\ell}\},
\]
we imply $X\times Y = \bigcup_{i'\in I'}W_i$ and $I'$ is finite.
\end{enumerate}
\end{proof}
\begin{theorem}\label{The:5:3}
Suppose that $X$ is compact, $Y$ is Hausdorff, $f:X\to Y$ is continuous, bijective, then $f$ is a \emph{homeomorphism}.
\end{theorem}
\begin{proof}
It suffices to show $f^{-1}$ is continuous.
Therefore, it suffices to show $(f^{-1})^{-1}(V)$ is closed, given that $V$ is closed in $X$:

Let $V\subseteq X$ be closed. Then $V$ is compact, which implies $f(V)$ is compact. 
Since $f(V)\subseteq Y$ is Hausdorff, we imply $f(V)$ is compact, i.e., $f(V)$ is closed.
\end{proof}


















