\input{book_command}
\usepackage{pdfpages}
\includeonly{week1/Monday}
\begin{document}
%Hello
\frontmatter
%\includepdf[pages={1}]{book_cover/cover/FrontCover}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Setting up title pages, type in the appropriate names here:

\booktitle{	\huge A Brief On \\
Functional Analysis}

\subtitle{MAT4010 Notebook}

\AuAff{
Instructor\\
Prof. Dong Wang\\ The Chinese University of Hongkong, Shenzhen}
\AuAff{
Tex Written By\\
Mr. Jie Wang\\ Visiting Student from Georgia Institute of Technology}


%% Print Half Title and Title Page:
\halftitlepage
\titlepage

\tableofcontents

%\listoffigures %optional
%\listoftables  %optional




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Optional Acknowledgments:

\acknowledgments
This book is taken notes from the MAT4010 in spring semester, 2021. These lecture notes
were taken and compiled in LATEX by Jie Wang, a visiting first-year Ph.D. student from Gatech.
The instructor has not edited this document. Students taking this course may use
the notes as part of their reading and reference materials. This version of the lecture
notes were revised and extended for many times, but may still contain many mistakes
and typos, including English grammatical and spelling errors, in the notes. It would
be greatly appreciated if those students, who will use the notes as their reading or
reference material, tell any mistakes and typos to Jie Wang for improving this notebook.
\authorinitials{Jie, 2021-01-11}  


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Optional notations:
\begin{notations}
\acro{$\mathbb{R}^n$}{$n$-dimensional real space}
\acro{$\mathbb{C}^n$}{$n$-dimensional complex space}
\acro{$\mathbb{R}^{m\times n}$}{set of all $m\times n$ real-valued matrices}
\acro{$\mathbb{C}^{m\times n}$}{set of all $m\times n$ complex-valued matrices}
\acro{$x_i$}{$i$th entry of column vector $\bm x$}
\acro{$a_{ij}$}{$(i,j)$th entry of matrix $\bm A$}
\acro{$\bm a_i$}{$i$th column of matrix $\bm A$}
\acro{$\bm a_i\trans$}{$i$th row of matrix $\bm A$}
\acro{$\mathbb{S}^n$}{set of all $n\times n$ real symmetric matrices, i.e., $\bm A\in\mathbb{R}^{n\times n}$ and $a_{ij}=a_{ji}$ for all $i,j$}
\acro{$\mathbb{H}^n$}{set of all $n\times n$ complex Hermitian matrices, i.e., $\bm A\in\mathbb{C}^{n\times n}$ and $\bar{a}_{ij}=a_{ji}$ for all $i,j$}
\acro{$\bm A\trans$}{transpose of $\bm A$, i.e, $\bm B=\bm A\trans$ means $b_{ji}=a_{ij}$ for all $i,j$}
\acro{$\bm A\Her$}{Hermitian transpose of $\bm A$, i.e, $\bm B=\bm A\Her$ means $b_{ji}=\bar{a}_{ij}$ for all $i,j$}
\acro{$\trace(\bm A)$}{sum of diagonal entries of square matrix $\bm A$}
\acro{$\bm 1$}{A vector with all $1$ entries}
\acro{$\bm 0$}{either a vector of all
zeros, or a matrix of all zeros}
\acro{$\bm e_i$}{a unit vector with the nonzero element at the $i$th entry}
\acro{$\mathcal{C}(\bm A)$}{the column space of $\bm A$}
\acro{$\mathcal{R}(\bm A)$}{the row space of $\bm A$}
\acro{$\mathcal{N}(\bm A)$}{the null space of $\bm A$}
\acro{$\Proj_{\mathcal{M}}(\bm A)$}{the projection of $\bm A$ onto the set $\mathcal{M}$}
\end{notations}
\mainmatter
\setcounter{page}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\include{week1/Monday}
%\include{week1/Wednesday}
%\include{week2/Monday}
%\include{week2/Wednesday}
%\include{week3/Monday}
%\include{week3/Wednesday}
%\include{week4/Monday}
%\include{week4/Wednesday}
%\include{week5/Monday}
%\include{week5/Wednesday}
%\include{week6/Monday}
%\include{week6/Wednesday}
%\include{week7/Monday}
%\include{week7/Wednesday}
%\include{week8/Monday}
%\include{week8/Wednesday}
%\include{week9/Monday}
%\include{week9/Wednesday}
%\include{week10/Monday}
%\include{week10/Wednesday}
%\include{week11/Monday}
%\include{week11/Wednesday}
%\include{week12/Monday}
%\include{week12/Wednesday}
%\include{week13/Monday}
%\include{week13/Wednesday}
%\include{week14/Monday}
%\include{week14/Wednesday}



\backmatter
\latexprintindex
\end{document} 