\section{Wednesday}

\begin{proposition}
$G$ is hamiltionian implies that for any nonempty $S\subseteq V$, $G-S$ has at most $|S|$ components.
\end{proposition}
\begin{theorem}
$G$ is Hamiltionian if for any vetices $v,w$ such that $(v,w)\notin E$,
\[
\mbox{deg}(v)+\mbox{deg}(w)\ge n
\]
\end{theorem}


\paragraph{Knapsacle Problem}
\[
\begin{array}{ll}
\max&\sum_{i=1}^nv_iu_i\\
\mbox{such that}&\sum_{i=1}^nw_iu_i\le W\\
&u_i\in\{0,1\},\ i=1,\dots,n
\end{array}
\]

This is the binary integer programming problem, which is NP-complete. We cannot find exact solution to this problem. There are $2^n$ possible solutions, which requires exponential time.

We can find some ``pseudo''-polynomial algorithm. Assumption:
\begin{enumerate}
\item
$W$ is an integer. 

\end{enumerate}

State: remaining capacity, define as $X_k$.

Stage: iterm $1,\dots,n$.
\[
J_N(x_N)=\left\{
\begin{aligned}
v_N &\quad\text{if }N\le X_N\\
0,&\quad\text{if }w_N>X_N
\end{aligned}
\right.
\]
\[
J_N(x_{N-1})
=
\left\{
\begin{aligned}
0+J_N(X_N),&\quad\text{if }w_{N-1}>X_{N-1}, \text{Here }X_N=X_{N-1}\\
\max\{v_{N-1}+J_N(x_N), 0+J_N(x_N)\}&,\text{ if }w_{N-1}\le X_{N-1}
\end{aligned}
\right.
\]

Here why the DP has ``pesduo''-polynomial time, and the $w_i$ should be integer?

Let's list states $\{x_1,\dots,x_N\}$. For each stage $k$, there would be $W+1$.

Each iteration at most $2$ computation, and therefore we face $2(W+1)\cdot N$

\subsection{Label Correcting Methods}
Shortest Path. Back up some information.

The label refers to the intermediate information.

$A^*$-algorithm; Bellman-Ford Algorithm

Benchmark: MILP, CPLEX, CVX.

\subsection{DP problems with perfect state information}

\paragraph{Linear Quadratic System}
Let $x_k\in\mathbb{R}^1$ be the state; $u_k\in\mathbb{R}^n$ be the control; $\omega_k\in\mathbb{R}^n$ be the disturbance.

System Dynamics.
\[
x_{k+1}=A_kx_k+B_ku_k+\omega_k
\]
Stage cost:
\[
x_k'Q_kx_k+u_k'R_ku_k
\]
Here $Q_k$ is required to be positive-semi-definite symmetric matrix; $R_k$ to be positive-definite symmetric matrix.
\[
J_{n-1}(x_{n-1})=\min_{u_{n-1}}\mathbb{E}_{\omega}
\left\{
x_{n-1}'Q_{n-1}x_{n-1}+u_{n-1}'R_{n-1}u_{n-1}
+
J_n(x_n)
\right\}
\]
where
\[
J_n(x_n)=(A_{n-1}x_{n-1}+Bu_{n-1}+\omega)'Q_n
(A_{n-1}x_{n-1}+Bu_{n-1}+\omega)
\]


\subsection{Linear Quadratic}
Dynamics:
\[
x_{k+1}=A_kx_k+B_ku_k+\omega_k
\]
Stage cost:
\[
x_k'Q_kx_k+u_k'R_ku_k, \ Q_k\succeq0, R_k\succ 0
\]
Cost to go function:
\[
J_k(x_k)=\min_{u_k}\mathbb{E}_{\omega_k}
[
x_k'Q_kx_k+u_k'R_ku_k
+
J_{k+1}(x_{k+1})
]
\]
The final cost is 
\[
J_N(x_N)=x_N'Q_Nx_N
\]

At $N-1$ stage
\begin{align*}
&x_{N-1}'Q_{N-1}x_N+u_{N-1}'R_{N-1}u_{N-1}\\
&+
\mathbb{E}
[
(A_{N-1}x_{N-1}+B_{N-1}u_{N-1}+\omega_{N-1})'
Q_{N}
(A_{N-1}x_{N-1}+B_{N-1}u_{N-1}+\omega_{N-1})
]
\end{align*}

Optimization model:
\[
\frac{\partial}{\partial u}(u'Hu+2r'u+c) = 2Hu+2r\implies
Hu = -r\implies
u^*=-H^{-1}r
\]

For $N-1$ stage, we have
\begin{align*}
H&=R_{N-1}+B_{N-1}'Q_NB_{N-1}\\
r'&=x_{N-1}'A_{N-1}'Q_NB_N\\
c&=x_{N-1}'(Q_{N-1}+A_{N-1}'Q_NA_{N-1})x_{N-1}
+
\mathbb{E}_{\omega}(\omega_{N-1}'Q_N\omega_{N-1})
\end{align*}

Therefore,
\[
u_{N-1}^*=-(R_{N-1}+B_{N-1}'Q_NB_{N-1})^{-1}B_NQ_NA_{N-1}x_{N-1},
\]
which is linaer in terms of $x_{N-1}$, which is so called linear controller.

Therefore,
\[
J_{N-1}(x_{N-1})
=
x_{N-1}'K_{N-1}x_{N-1}
+
\mathbb{E}(\omega'_{N-1}Q_N\omega_{N-1})
\]
The linear controller will be tested during mid-term or final.

\[
J_0(x_0)=x_0'K_0x_0+\sum_{k=0}^{N-1}\mathbb{E}_\omega\{\omega_k'K_{k+1}\omega_k\}
\]
\[
K_{k-1}=f(K_k)\implies K_{k-1}=K_k,\ \text{ for large }k.
\]

\paragraph{Riccati Equation}

for stationary ststem, i.e., $A_k=A,B_k=B,Q_k=Q,R_k=R$.
\[
K=A'(K - KB(B'KB+R)^{-1}B'K)A+Q
\]

Therefore as time goes long enough, the cost to go function $J_k(x_k)$ will be a constant plus over stages. such a constant $K$ is called the \emph{optimal stationary controller}.

Stability. For $u^*=Lx$, we imply
\[
x_{k+1}=Ax_k+Bu_k+\omega_k=(A+BL)x_k+\omega_k
\]
Care about $\lim_{k\to\infty}(A+BL)^k=0$.

Here
\[
L = -(B'KB+R)^{-1}B'KA
\]
It suffices to solve
\[
P_{k+1} = A^2\left(
P_k-\frac{B^2P_k^2}{B^2P_k+R}
\right)
+
Q
\]

Then consider the case that $A_k,B_k$ are all random matrices, i.e., independent. $Q_k\succeq0, R_k\succ0$.

\[
L_k = -\left(
R_k+\mathbb{E}(B_k'K_{k+1}B_k)
\right)^{-1}
\mathbb{E}(B_j'K_{k+1}A_k)
\]

Note that
\[
P_{\infty}
=
\frac{\mathbb{E}A^2 RP}{R+EB^2P}
+
\frac{\mathbb{E}A^2\mathbb{E}B^2 - (\mathbb{E}A)^2 (\mathbb{E}B)^2}{R + \mathbb{E}B^2P}
\]

Certainty Equivalence

State: $x_k$, inventory level

Control: $u_k\ge0$, number of orders placed

Disturbance: $\omega_k$: damand

Dynamics:
\[
x_{k+1 } = (x_k+u_k-\omega_k)
\]

Stage cost:
\begin{itemize}
\item
Ordering cost: $c\cdot u_k$
\item
Maintaining cost: full backlog.$ \mathbb{E}_{\omega_k}(r(x_k+u_k-\omega_k)).$

where $r(z) =hz^+ + pz^-$ is a convex function.

We imply the maintaining cost is $H(x_k+u_k)$, which is convex.
\end{itemize}

\[
J_k(x_k) = \min_{u_k\ge0}
\left\{
cu_k+H(x_k+u_k)
+
\mathbb{E}[J_{k+1}(x_k+u_k-\omega_k)]
\right\}
\]

Define $Y=x_k+u_k$, and therefore
\[
G(Y) = cY+H(Y)+\mathbb{E}[J(Y-\omega_k)]
\]

Therefore, we can always set $Y=x_k+u_k=S_k$, where $S_k$ minimizes $G(Y)$.









