
\chapter{Week3}

\section{Tuesday}\index{week3_Tuesday_lecture}
\subsection{Introduction}
\subsubsection{Motivation of Linear Algebra}
So, we raise the question again, why do we learn LA?
\begin{itemize}
\item
Baisis of AI/ML/SP/etc.\\
In information age, \textit{artificial intelligence, machine learning, structured programming}, and otherwise gains great popularity among researchers. LA is the basis of them, so in order to explore science in modern age, you should learn LA well.
\item
Solving linear system of equations.\\
How to solve linear system of equations efficiently and correctly is the \emph{key} question for mathematicians.
\item
Internal grace.\\
LA is very beautiful, hope you enjoy the beauty of math.
\item
Interview questions.\\
LA is often used for interview questions for phd. The interviewer usually ask difficult questions about LA.
\end{itemize}
\subsubsection{Preview of LA}
The main branches of Mathematics are given below:
\[
\text{mathematics}\begin{cases}
\mbox{Analysis + Calculus} \\
\mbox{Algebra: foucs on structure} \\
\mbox{Geometry}
\end{cases}
\]
All parts of math are based on \emph{axiom systems}. And \emph{LA} is the significant part of \textit{Algebra}, which focus on the linear structure.

\subsection{Review of 2 weeks}
\paragraph{How to solve linear system equations?}
The basic method is \emph{Gaussian Elimination}, and the main idea is \textit{induction} to make simpler equations.
\begin{itemize}
\item
Given one equation $ax=b$, we can easily sovle it:
\[
\begin{array}{ll}
\mbox{If $a=0$, there is no solution}
&
\mbox{otherwise $x=\frac{b}{a}$}.
\end{array}
\]
\item
We could solve $1\times 1$ system. By induction, if we could solve $n\x n$ systems, then we can solve $(n+1)\x(n+1)$ systems.
\end{itemize}

In the above process, math notations is needed:
\begin{itemize}
\item
matrix multiplication
\item
matrix inverse
\item
transpose, symmetric matrices
\end{itemize}
So in first two weeks, we just learn two things:
\begin{itemize}
\item
\textit{linear system could be solved \emph{almost} by G.E.}
\item
\textit{Furthermore, Gaussian Elimination is (almost) LU decomposition.}
\end{itemize}
But there is a question remained to be solved:
\paragraph{How to solve linear singular system equations?}
\begin{itemize}
\item
When does the system have no solution, when does the system have infinitely many solutions? (Note that singular system don't has unique solution.)
\item
If it has infinitely many solutions, how to find and express these solutions?
\end{itemize}
If we express system into matrix form, the question turns into:
\begin{quotation}
How to solve the rectangular?
\end{quotation}

\subsection{Examples of solving equations}

\begin{itemize}
\item
For square case, we often convert the system into $\bm{Ux} = \bm c$, where $\bm U$ is of \textit{row echelon form}.
\item
However, for rectangular case, \textit{row echelon form}(ref) is not enough, we must convert it into \emph{reduced row echelon form}(rref):
\[\bm U\text{(ref)} = \left[
\begin{array}{@{}ccccccc@{}}
\cellcolor{cyan!50}1&0&\x&\x&\x&0&\x\\
0&\cellcolor{cyan!50}1&\x&\x&\x&0&\x\\
0&0&0&0&0&\cellcolor{cyan!50}1&\x\\
0&0&0&0&0&0&0
\end{array}
\right]
\implies
\bm R\text{(rref)} = \left[
\begin{array}{@{}ccccccc@{}}
\cellcolor{cyan!50}1&0&\x&\x&\x&0&\x\\
0&\cellcolor{cyan!50}1&\x&\x&\x&0&\x\\
0&0&0&0&0&\cellcolor{cyan!50}1&\x\\
0&0&0&0&0&0&0
\end{array}
\right]\]
\end{itemize}
\begin{example}
We discuss how to solve \emph{square} matrix of \emph{rref}:
\begin{itemize}
\item
If all rows have nonzero entry, we have:
\[
\begin{bmatrix}
1&&\bigzero&\\
&1&&\\
&&1&\\
&\bigzero&&1
\end{bmatrix}\bm x = \bm c
\implies \bm x = \bm c
\]
\item
But note that \textit{some rows could be all zero}:
\[
\begin{bmatrix}
1&&&\\
&1&&\\
&&1&\\
&&&0
\end{bmatrix}\bm x = \bm c\implies
\begin{cases}
x_1 = c_1\\
x_2 = c_2\\
x_3 = c_3\\
0 = c_4
\end{cases}
\]
So the solution results have two cases:
\begin{itemize}
\item
If $c_4\ne0$, we have no solution of this system.
\item
If $c_4=0$, we have infinitely many solutions, which can be expressed as:
\[
x_{\text{complete}} = \begin{pmatrix}
c_1\\c_2\\c_3\\x_4
\end{pmatrix} = \begin{pmatrix}
c_1\\c_2\\c_3\\0
\end{pmatrix}+x_4\begin{pmatrix}
0\\0\\0\\1
\end{pmatrix}
\]
where $x_4$ could be arbitarary number.
\end{itemize}
\end{itemize}
Hence, for square system, does Gaussian Elimination work?

Answer: Almost, except for the ``pivot=0''case:
\begin{itemize}
\item
All pivots $\ne0\implies$ the system has unique solution.
\item
Some pivots $=0$ (The matrix is singular)
\begin{enumerate}
\item
No solution. (When LHS $\ne$ RHS)
\item
Infinitely many solutions.
\end{enumerate}
\end{itemize}
\end{example}

\subsubsection{
Review of G.E. for Nonsingular case
}
We use matrix to represent system of equations: 
\[
\begin{cases}
a_{11}x_1 + a_{12}x_2 + \dots + a_{1n}x_n = b_1 \\ 
a_{21}x_1 + a_{22}x_2 + \dots + a_{23}x_n = b_2  \\
\dots 	 	\\
a_{m1}x_1 + a_{m2}x_2 + \dots + a_{m3}x_n = b_m  
\end{cases}
\implies
\bm{Ax} = \bm b\]
By postmultiplying $\bm E_{ij}$ or $\bm P_{ij}$, we are essentially doing one step of elimination:
\[
\begin{array}{lll}
\bm E_{ij}\bm A\bm x = \bm E_{ij}\bm b
&
\mbox{or}
&
\bm P_{ij}\bm A\bm x = \bm E_{ij}\bm b
\end{array}
\]
By several steps of elimination, we obtain the final result:
\[
\bm{\hat LPAx} = \bm{\hat LP}\bm b
\]
where $\bm{\hat LPA}$ represents an upper triangular matrix $\bm U$, $\hat L$ is the lower triangular matrix.

 Equivalently, we obtain
\[
\bm{\hat LPA} = \bm U\implies
\bm{PA} = \bm{\hat L}^{-1}\bm U\triangleq\bm{LU}
\]
Hence, Gaussian Elimination is almost the $LU$ decomposition.

\subsubsection{Example for solving rectangular system of rref}
Recall the definition for rref:
\begin{definition}[reduced row echelon form]
Suppose a matrix has $r$ \textit{nonzero} rows, each row has leading $1$ as pivots. If all columns with pivots (call it pivot column) are all zero entries apart from the pivot in this column, then this matrix is said to be \emph{reduced row echelon form}(\emph{rref}).
\end{definition}
Next, we want to show how to solve a rectangular system of rref. Note that in last lecture we study the solution to a rectangular system  is given by:
\[
\bm x_{\text{complete}} =\bm x_p +\bm x_{\text{special}}.
\]

\begin{example}
Solve the system 
\[
\begin{bmatrix}
1&3&0&0\\0&0&1&1\\0&0&0&0
\end{bmatrix}\bm x = \bm c.
\]
\paragraph{Step 1: Find null space} Firstly we solve for $\bm{Rx} = \bm 0$:
\[
\begin{bmatrix}
1&3&0&0\\0&0&1&1\\0&0&0&0
\end{bmatrix}\begin{bmatrix}
x_1\\x_2\\x_3\\x_4
\end{bmatrix} = \begin{bmatrix}
0\\0\\0
\end{bmatrix}\implies \left\{\begin{rgathered}
x_1+3x_2=0\\
x_3+x_4=0
\end{rgathered}\right.
\]

Then we express the \emph{pivot variables} in the form of \emph{free variables.}
\begin{quotation}
Note that the pivot columns in $\bm R$ are column $1$ and $3$, so the pivot variable is $x_1$ and $x_3$. The free variable is the remaining variable, say, $x_2$ and $x_4$.
\end{quotation}
The expressions for $x_1$ and $x_3$ are given by:
\[
\left\{\begin{lgathered}
x_1=-3x_2\\
x_3=-x_4
\end{lgathered}\right.
\]

Hence, all solutions to $\bm{Rx} = \bm 0$ are
\[
\bm x_{\text{special}} = \begin{bmatrix}
-3x_2\\x_2\\-x_4\\x_4
\end{bmatrix} = x_2\begin{bmatrix}
-3\\1\\0\\0
\end{bmatrix} + x_4\begin{bmatrix}
0\\0\\-1\\1
\end{bmatrix}
\]
where $x_2$ and $x_4$ can be taken arbitararily.
\paragraph{Step 2: Find one particular solution to $\bm{Rx} = \bm c$} The trick for this step is to set $x_2=x_4=0$. (\textit{set free variable to be zero and then derive the pivot variable.}):
\[
\begin{bmatrix}
1&3&0&0\\0&0&1&1\\0&0&0&0
\end{bmatrix}\begin{bmatrix}
x_1\\0\\x_3\\0
\end{bmatrix} = \begin{bmatrix}
c_1\\c_2\\c_3
\end{bmatrix}
\implies
\left\{\begin{rgathered}
x_1 =c_1\\
x_3 =c_2\\
0 =c_3
\end{rgathered}\right.
\]
which follows that:
\begin{itemize}
\item
if $c_3 = 0$, then exists particular solution $\bm x_p = \begin{bmatrix}
c_1\\0\\c_2\\0
\end{bmatrix}$;
\item
if $c_3\ne 0$, then $\bm{Rx} = \bm c$ has no solution.
\end{itemize}
\paragraph{Final solution} If assume $c_3= 0$, then all solutions to $\bm{Rx} = \bm c$ are given by:
\[
\bm x_{complete} = \bm x_p + \bm x_{\text{special}} = 
\begin{bmatrix}
c_1\\0\\c_2\\0
\end{bmatrix} + x_2\begin{bmatrix}
-3\\1\\0\\0
\end{bmatrix} + x_4\begin{bmatrix}
0\\0\\-1\\1
\end{bmatrix}
\]
\end{example}
Next we show how to solve a general rectangular:
\newpage
\subsection{How to solve a general rectangular}
For linear system $\bm{Ax} = \bm b$, where $\bm A$ is rectangular, we can solve this system as follows:
\paragraph{Step 1: Gaussian Elimination}
With proper row permutaion (postmultiply $\bm P_{ij}$) and row transformation (postmultiply $\bm E_{ij}$), we convert $\bm A$ into $\bm R$(rref), then we only need to solve $\bm{Rx} = \bm c$.
\begin{example}
The first example is a $3\x 4$ matrix with two pivots:
\[
\bm A = \begin{bmatrix}
1&1&2&3\\2&2&8&10\\3&3&10&13
\end{bmatrix}
\]
Clearly $a_{11}=1$ is the first pivot, then we clear row 2 and row 3 of this matrix:
\[
\bm A
\xLongrightarrow[\bm E_{31} = \begin{bmatrix}
1&0&0\\0&1&0\\-3&0&1
\end{bmatrix}]{\bm E_{21} = \begin{bmatrix}
1&0&0\\-2&1&0\\0&0&1
\end{bmatrix}}
\begin{bmatrix}
1&1&2&3\\0&0&4&4\\0&0&4&4
\end{bmatrix}
\xLongrightarrow[\bm E_{32} = \begin{bmatrix}
1&0&0\\0&1&0\\0&-1&1
\end{bmatrix}]{\bm E_{12} = \begin{bmatrix}
1&-\frac{1}{2}&0\\0&1&0\\0&0&1
\end{bmatrix}}
\begin{bmatrix}
1&1&0&1\\0&0&4&4\\0&0&0&0
\end{bmatrix}\\
\]
\[\implies \begin{bmatrix}
1&1&0&1\\0&0&1&1\\0&0&0&0
\end{bmatrix}\]
If we want to solve $\bm{Ax} = \bm b$, firstly we should convert $\bm A$ into $\begin{bmatrix}
1&1&0&1\\0&0&1&1\\0&0&0&0
\end{bmatrix}$(rref).
\end{example}

Then we should identify \emph{pivot variables} and \emph{free variables}. we can follow the proceed below:
\[
\mbox{pivots}
\implies
\mbox{pivot columns}
\implies
\mbox{pivot variables}
\]
\begin{example}
we want to identify \emph{pivot variables} and \emph{free variables} of $\bm R$: 
\[
\bm R = \left[
\begin{array}{@{}ccccccc@{}}
\cellcolor{cyan!50}1&0&\x&\x&\x&0&\x\\
0&\cellcolor{cyan!50}1&\x&\x&\x&0&\x\\
0&0&0&0&0&\cellcolor{cyan!50}1&\x\\
0&0&0&0&0&0&0
\end{array}
\right]
\]
The pivot are $r_{11},r_{22},r_{36}$. So the pivot columns are column $1,2,6$. So the \textit{pivot variables} are $x_1,x_2,x_6$; the \textit{free variables} are $x_3,x_4,x_5,x_7$.
\end{example}
\paragraph{Step2: Compute null space $N(\bm A)$}
In order to find $N(\bm A)$, it suffices to compute $N(\bm R)$. The space $N(\bm R)$ has $(n-r)$ dimensions, so it suffices to get $(n-r)$ special solutions first:
\begin{itemize}
\item
For each of the $(n-r)$ free variables,
\begin{itemize}
\item
set the value of \emph{it} to be $1$;
\item
set the value of other \emph{free variables} to be 0;
\item
Then solve $\bm{Rx} = \bm 0$ (to get the value of pivot variables) to get the special solution.
\end{itemize}
\begin{example}
Continue with $3\x 4$ matrix example:
\[
\bm R = \begin{bmatrix}
1&1&0&1\\0&0&1&1\\0&0&0&0
\end{bmatrix}\]
We want to find special solutions to $\bm{Rx} = \bm 0$:
\begin{enumerate}
\item
Set $x_2=1$ and $x_4=0$. Solve $\bm{Rx} = \bm 0$, then $x_1=-1$ and $x_3 = 0$. \\Hence one special solution is $y_1 = \begin{bmatrix}
-1\\1\\0\\0
\end{bmatrix}$.
\item
Set $x_2=0$ and $x_4=1$. Solve $\bm{Rx} = \bm 0$, then $x_1=-1$ and $x_3 = -1$. \\Then another special solution is $y_2 = \begin{bmatrix}
-1\\0\\-1\\1
\end{bmatrix}$.
\end{enumerate}
\end{example}
\item
Then $N(\bm A)$ is the collection of linear combinations of these special solutions:
\[
N(\bm A) = \Span(y_1,y_2,\dots,y_{n-r}).
\]
\begin{example}
We continue the example above, when we get all special solutions 
\[
\begin{array}{ll}
y_1 = \begin{bmatrix}
-1\\1\\0\\0
\end{bmatrix}
&
y_2 = \begin{bmatrix}
-1\\0\\-1\\1
\end{bmatrix},
\end{array}
\] 
\emph{the null space contains all linear combinations of the special solutions}:
\[
\bm x_{\text{special}} = \Span(\begin{bmatrix}
-1\\1\\0\\0
\end{bmatrix},\begin{bmatrix}
-1\\0\\-1\\1
\end{bmatrix}) = x_2\begin{bmatrix}
-1\\1\\0\\0
\end{bmatrix} + x_4\begin{bmatrix}
-1\\0\\-1\\1
\end{bmatrix}
\]
where $x_2,x_4$ here could be arbitarary.
\end{example}
\end{itemize}

\paragraph{Step3: Compute a particular solution $\bm x_p$}
The easiest way is to ``read'' from $\bm{Rx} = \bm c$:
\begin{quotation}
\begin{itemize}
\item
\emph{Guarantee the existence of the solution}. Suppose $\bm R\in\mathbb{R}^{m\x n}$ has $r (\le m)$ pivot variables, then it has $(m-r)$ zero rows and $(n-r)$ free variables. For the existence of solutions, the value of entries of $\bm c$ which correspond to zero rows in $\bm R$ must also be zero.
\begin{example}
If $\bm{Rx} = \begin{bmatrix}
1&3&0&2\\0&0&1&4\\\bm 0&\bm 0&\bm 0&\bm 0
\end{bmatrix}\bm x = \bm c = \begin{bmatrix}
c_1\\c_2\\c_3
\end{bmatrix}$, then in order to have a solution, we must let $c_3\ne 0$.
\end{example}
\item
If the condition above is not satisfied, then the system has no solution. Let's preassume the satisfaction of such a condition. To compute a particular solution $\bm x_p$, we set \textit{the value for all \emph{free variables} of $\bm x_p$ to be zero, and the value for the pivot variables are from $\bm c$.} 

More specifically, the first entry in $\bm c$ is exactly the value for the first pivot variable;the second entry in $\bm c$ is exactly the value for the second pivot variable$\dots\dots$, and the remaining entries of $\bm x_p$ are set to be zero.
\begin{example}
If $\bm{Rx} = \begin{bmatrix}
1&3&0&2\\0&0&1&4\\\bm 0&\bm 0&\bm 0&\bm 0
\end{bmatrix}\bm x = \bm c = \begin{bmatrix}
c_1\\c_2\\0
\end{bmatrix}$, we want to compute particular solution 
\[\bm x_p = \begin{bmatrix}
x_1\\x_2\\x_3\\x_4
\end{bmatrix}\]
As we know $x_2,x_4$ are free variable, $x_2=x_4=0$; and $x_1,x_3$ are pivot variable, so we have $\begin{pmatrix}
x_1\\x_3
\end{pmatrix} = \begin{pmatrix}
c_1\\c_2
\end{pmatrix}$.\\
Hence the solution for $\bm{Rx} = \bm c$ is 
\[
\bm x_p=
\begin{bmatrix}
c_1\\0\\c_2\\0
\end{bmatrix}.
\]
\end{example}
\end{itemize}
\end{quotation}
\paragraph{Final step: Obtain complete solutions}
All solution of $\bm{Ax} = \bm b$ are 
\[
\bm x_{\text{complete}} = \bm x_p + \bm x_{\text{special}},
\] 
where $x_{\text{special}}\in N(\bm A)$. Note that $\bm x_p$ is defined in step3, $\bm x_{\text{special}}$ is defined in step2.

However, where does the number $r$ come? $r$ denotes the \emph{rank} of a matrix, which will be discussed in the next lecture.



















